Smooth and Persistent Forecasts of German GDP

### 2025-06-20 HWI SSA_27_05_2025_with new poster/05.06.2025/paper.tex
up

\usepackage{graphicx} % Required for inserting images
\usepackage{tabularx}
\usepackage{natbib}
\usepackage[spaces,hyphens]{url}
\usepackage[hidelinks]{hyperref}
\usepackage{setspace}
\usepackage{booktabs}
\usepackage{hyphenat}
\usepackage{color}
%\usepackage{subfig}
% \usepackage[justification=left,hang]{caption}
\usepackage[justification=raggedright,hang]{caption}
\usepackage{subcaption}
	%\captionsetup[subfigure]{labelfont=normal,textfont=normalfont,singlelinecheck=off,justification=centering}
\usepackage[table]{xcolor} 
\usepackage{colortbl}
\usepackage[hang]{footmisc}
\usepackage{appendix}

@@

\begin{abstract} %max.
This paper addresses the trade-off between informational efficiency and forecast smoothness in economic forecasting. Efficient forecasts are often volatile, limiting their practical usefulness. Building on the Smooth Sign Accuracy (SSA) framework, we propose a novel method that integrates Multivariate Smooth Sign Accuracy (M-SSA) to extract smoothed components from leading indicators. This approach enhances the signal-to-noise ratio and allows for systematic control over forecast volatility and timing. Applied to German GDP, our method yields smoothed forecasts that match or exceed the accuracy of traditional models, particularly over medium-term horizons. While smoother forecasts may lag slightly, this can be mitigated by adjusting the forecast horizon. The findings offer practical tools for forecasters and policymakers operating under adjustment constraints.

@@

\begin{quote}
     \quad As we parse the incoming information, we are focused on separating the signal from the noise as the outlook evolves.\\
    \textit{Speech by Federal Reserve Board Chair Jerome H. Powell, March 07, 2025, \\
    University of Chicago Booth School of Business}
\end{quote}

@@

This study proposes a methodological innovation that extends the classical direct forecasting framework by integrating dimension-reduced information extracted through Multivariate Smooth Sign Accuracy (M-SSA). The core idea is to replace raw
%, unprocessed 
leading indicators with smoothed and filtered components derived from the M-SSA procedure, thereby improving the signal-to-noise ratio in the predictor variables used to forecast GDP. We apply this framework to well-established German economic indicators to show its ability to improve forecasts of GDP and GDP trends, especially over mid-term horizons ranging from two to four quarters ahead, using a variety of sample periods and performance metrics.

Surprisingly, we find that smoothing indicator-based forecasts often provides useful leading information for GDP. We also show that while smoother GDP forecasts lag efficient MSE forecasts slightly, this can be mitigated by adjustments in the forecast horizon.\footnote{The lag reflects the trilemma highlighted by \cite{McElroy2019} in which forecasters choose between forecast accuracy, timeliness and volatility.}

@@

Section \ref{sec:data} describe our data series, the sample period used, and documents their dynamic behaviour. Section \ref{sec:direct_forecast} provides benchmark forecasts using simple direct projections of GDP based on our indicators. It then demonstrates how in-sample forecast performance may improve when we ``de-noise'' the indicators prior to projection with simple smoothing filters. Section \ref{sec:mSSA} defines the M-SSA method and shows how its forecast performance compares to that of direct projections as well as how it varies with forecast horizon. Section \ref{sec:mSSA_component} considers the ability of our indicators to forecast German GDP growth using the M-SSA framework, and Section \ref{sec:revisions} investigates the temporal stability of the estimated forecasting equations. Section \ref{sec:conclusions} summarizes the results and concludes. The Appendix provides some additional results.

\section{Data and Dependence}\label{sec:data}
\subsection{Data}
The analysis employs a set of well-established leading indicators to forecast quarterly German GDP \citep{Drechsel2012financial,Heinisch2018bottom}, including industrial production (IP), the ifo Business Climate Index (ifo\_c), the Economic Sentiment Indicator (ESI), and the term spread between 10-year bond yields and the 3-month EURIBOR rate (spre\_10y\_3m) covering the period 1995 to 2024. All series are used at a quarterly frequency, with a particular emphasis on forecast horizons of 2 or more quarters, reflecting medium-term forecasting objectives. 
To ensure comparability and mitigate scale effects, all indicators (except the spread, which is differenced only) are log-differenced and standardized, the latter applied solely for ease of visual inspection.
Additionally, extreme outliers associated with the COVID-19 pandemic are trimmed to prevent distortion of the underlying signal.
We take into account the publication lag of the indicators and the GDP series.\footnote{Data revisions are not considered, given that previous studies have shown that their overall effect on forecast performance is minor.} 
While GDP is available with a one-quarter delay, and IP is released with a two-month lag, business surveys and financial data are published without any lag. All series are therefore re-aligned to reflect the forecasters' available information, with GDP and IP shifted forward by their respective lags. Tables \ref{tab:data_stucture} and \ref{tab:data_stucture_aligned} show the data structure, as published by mid-February 2025.

%\begin{table}[ht]
%\centering
%\begin{tabular}{rrrrrr}
%  \hline
% & GDP & ip & ifo\_c & ESI & spr\_10y\_3m \\ 
%  \hline
%2024-07-01 & & 90.700 & 85.800 & 92.500 & -1.300 \\ 
%  2024-08-01 &  & 93.200 & 84.900 & 90.900 & -1.400 \\ 
%  2024-09-01 &902.571   & 91.200 & 83.700 & 89.800 & -1.400 \\ 
%  2024-10-01 &  & 90.800 & 84.100 & 90.700 & -1.200 \\ 
%  2024-11-01 &  & 92.200 & 83.800 & 89.300 & -0.900 \\ 
%  2024-12-01 & 900.761 &  & 82.800 & 86.900 & -0.800 \\ 
%  2025-01-01 &  &  & 82.500 & 88.100 & -0.200 \\ 
%   \hline
%\end{tabular}
%\caption{Ragged end.  } 
%\label{tab:data_stucture}
%\end{table}% latex table generated in R 4.2.2 by xtable 1.8-4 package
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
  \hline
 & GDP & ip & ifo\_c & ESI & spr\_10y\_3m \\ 
  \hline
2024-07-01 & 902.571 & 90.700 & 85.800 & 92.500 & -1.300 \\ 
  2024-08-01 &  & 93.200 & 84.900 & 90.900 & -1.400 \\ 
  2024-09-01 &  & 91.200 & 83.700 & 89.800 & -1.400 \\ 
  2024-10-01 & 900.761 & 90.800 & 84.100 & 90.700 & -1.200 \\ 
  2024-11-01 &  & 92.200 & 83.800 & 89.300 & -0.900 \\ 
  2024-12-01 &  &  & 82.800 & 86.900 & -0.800 \\ 
  2025-01-01 &  &  & 82.500 & 88.100 & -0.200 \\ 
   \hline
\end{tabular}
\caption{Ragged end.  } 
\label{tab:data_stucture}
\end{table}
% Thu May  8 05:39:20 2025
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
  \hline
 & GDP & ip & ifo\_c & ESI & spr\_10y\_3m \\ 
  \hline
2024-05-01 &  & 93.400 & 84.900 & 90.900 & -1.400 \\ 
  2024-08-01 &  & 90.700 & 83.700 & 89.800 & -1.400 \\ 
  2024-09-01 & 902.571 & 93.200 & 84.100 & 90.700 & -1.200 \\ 
  2024-10-01 &  & 91.200 & 83.800 & 89.300 & -0.900 \\ 
  2024-11-01 &  & 90.800 & 82.800 & 86.900 & -0.800 \\ 
  2024-12-01 & 900.761 & 92.200 & 82.500 & 88.100 & -0.200 \\ 
   \hline
\end{tabular}
\caption{Data aligned at sample end.} 
\label{tab:data_stucture_aligned}
\end{table}

 Monthly series are transformed to quarterly series and are displayed in Figure \ref{fig:data}, while Figure \ref{fig:data_lags} offers a close-up view of the financial crisis and the pandemic.\footnote{Monthly series are averaged to quarterly values.} The latter figure illustrates that the real-time GDP and industrial production series are synchronized at dips and peaks (coincident), whereas the other indicators tend to lead somewhat. The latter can be leveraged in a multivariate approach to enhance forecasting performance compared to univariate benchmarks.

% \begin{figure}[htbp]
% \begin{center}
% \includegraphics[height=3in, width=4.5in]{./Figures/Data.pdf}
% \caption{Data: standardized log-differences, trimmed to $\pm 3$.
% \label{data}}
% \end{center}
% \end{figure}

% \begin{figure}[htbp]
% \begin{center}
% \includegraphics[height=3in, width=4.5in]{./Figures/data_lags.pdf}
% \end{center}
% \caption{Data: leads and lags during financial crisis (left) and Pandemic (right).
% \label{data_lags}}
% \end{figure}

@@

%Hence, we are targeting a series that comprises the publication lag and the forecast horizon.

%[Simon and Katja: not sure if we need the following sentences:]
%Benchmark for analysis are the historical mean and the AR-predictions

%%%%%%%%%%%%%%%%%

%\footnote{Code is provided at \url{ }.}
%https://github.com/wiaidp/R-package-SSA-Predictor/blob/main/SSA%20Tutorials/tutorial_7.4_MSSA_component_predictor.R}}
%Among our results to date, we find that industrial production and economic outlook surveys provide much more valuable information than the yield curve. We also find that extreme events, such as the Great Financial Crisis or the COVID-19 pandemic, have an important impact on the forecasts. This reflects a common tradeoff facing forecasters: the choice between high accuracy in over the whole period and avoiding rare but large and costly forecast errors in extraordinary situations.

@@

Figures \ref{CCF} (full sample) and \ref{CCF_wc} (data prior to the pandemic) show the sample cross-correlation functions (CCF's) between current and lagged GDP and the two survey indicators. Both show strong correlations that peak at leads relative to GDP. Results for IP showed no lead, while correlations for the term spread were weaker. The Figures also show that the precise strengths of the correlations are influenced by the outliers associated with the COVID-19 pandemic. In the interest of model stability, we exclude data from Q4-2019 to Q4-2020 (pandemic) when modelling the dynamics for the multivariate filters developed below.

@@

We then estimated VARMA models to capture these dynamics, selecting a parsimonious VAR(1) specification.\footnote{We used the \texttt{refVARMA()} procedure in the \cite{MTSpackage} \texttt{MTS} package, which sets VARMA coefficients with t-statistics $< 1.5$ to zero. See \cite{tsay2013multivariate}. We intend to explore richer dynamic structures in subsequent work.} Figure \ref{ma_inv_BIP} shows the implied responses of GDP to shocks to each variable.\footnote{Shown are the MA coefficients for GDP from the vector moving-average (VMA) model obtained by inversion of the VAR(1) representation (i.e. the impluse-response functions, or IRFs, of GDP to the correlated reduced-form shocks.)}

\begin{figure}[h]
    \begin{center}
        \includegraphics[height=3in, width=6in]{./Figures/ma_inv_BIP.pdf}
        \caption{VMA Coefficients for GDP Implied by VAR(1)\\Estimation with pandemic (left) and without pandemic (right).
        \label{ma_inv_BIP}}\end{center}
\end{figure}

The model summarizes key features of the data after we account for publication lags. GDP responds most strongly to shocks to the two survey indicators (ifo\_c and ESI) after a modest lag. There also appears to be a more modest response to IP shocks after a slightly longer lag, but this vanishes once data after the start of the pandemic is included.\footnote{There is also weaker evidence of a response to the term spread (spr\_10y\_3m) at longer lags. This should not be confused with the implied response to a \textit{structural} shock monetary policy shock.} The delayed response of GDP to movements in these various indicators suggests that multivariate models may provide better GDP forecasts than univariate models, a question we pursue in the next section.
%, where lagging series (GDP and IP, after accounting for publication lags) depend on indicators leading in relative terms. In particular, the impulse response of the spread suggests the relevance of a univariate model for this specific leading indicator.

@@

\begin{equation}
    u_{t+h} = \beta \cdot \mathbf{X}_t + \varepsilon_t
    \label{eq:direct_forecast}
\end{equation}

where
\begin{itemize}
    \item $u_{t+h}$ is the growth rate of GDP, $h = \{0, 1, 2, ... \}$
    \item $\mathbf{X}_t \equiv [1,u_{t-1}, ip_t, ifo\_c_t, ESI_t, spr\_10y\_3m_t]'$ or their transformed counterparts and $u_{t-1}$ accounts for the publication lag of GDP.\footnote{Including lagged GDP enables the nesting of the AR forecast within the broader modeling framework.}
    \item $\hat{\beta}$ is estimated by least-squares, omitting observations affected by the pandemic.
\end{itemize}

@@

\begin{figure}[h]
    \begin{center}
        \includegraphics[width=0.9\textwidth]{./Figures/direct_wc_all.pdf}
        %\includegraphics[height=3in, width=4.5in]{./Figures/direct_wc_all.pdf}
        \caption{Direct forecasts (red lines) based on all indicators and full sample information (without pandemic): GDP (black line) shifted by zero, one, two, and three quarters.
        \label{direct_wc}}
    \end{center}
\end{figure}

Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} compare actual and fitted values for GDP at forecast horizons $h = 0, \ldots, 3$, using the four indicators and lagged GDP, without applying any filtering. Figure \ref{direct_wc} shows results over the sample without the pandemic, while \ref{direct_wc_financial_crisis} provides a closer look at the fit during the financial crisis. Table \ref{tab:pvaluedhp} shows p-values for tests of the joint null hypothesis that all coefficients on the indicators are zero (HAC-adjusted Wald test).

\begin{figure}[h]
    \begin{center}
        \includegraphics[width=0.9\textwidth]{./Figures/direct_wc_financial_crisis.pdf}
        \caption{Direct forecast across the financial crisis:  regression based on all indicators and full sample information, without pandemic.
        \label{direct_wc_financial_crisis}}
    \end{center}
\end{figure}

The results indicate that as the forecast horizon $h$ increases, the turning points (i.e., peaks and troughs) of the direct forecasts become increasingly right-shifted (lagged) relative to the target GDP. Additionally, the p-values of the Hac-adjusted Wald test precipitously increase from $h=2$ to $h=3$, see Table \ref{tab:pvaluedhp}, implying that the indicators become statistically insignificant for forecast horizons of more than two quarters.\footnote{ 
These findings are confirmed by the relative root mean-square errors comparing direct forecasts against the mean benchmark, see Table \ref{rRMSE_mSSA_direct_mean_without_covid8} in the appendix.}

@@

Figure \ref{hp_160} presents both classic two-sided and one-sided HP(160) (denoted HP-C for \textit{Concurrent}) filters (top panels) along with their associated amplitude functions in the frequency domain (bottom panels). It confirms that the HP-C filter assigns greater weight to yearly components—relevant in a mid-term forecast context— than the classic quarterly HP filter, while also reducing undesirable high-frequency noise. Therefore, while the choice of the HP(160) filter in this example
is somewhat arbitrary, it should serve the purpose of reducing high-frequency noise while preserving longer-duration movements.\footnote{A comprehensive technical analysis of the effect of $\lambda$ on the resulting GDP predictor can be conducted using the M-SSA package (\url{https://github.com/wiaidp/R-package-SSA-Predictor}).}

@@

Figures \ref{direct_hp_forecasts} and \ref{direct_hp_forecasts_financial_crisis} provide results comparable to those of Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} but now using indicators smoothed with the HP-C filter. As before, the first Figure covers the full sample without the pandemic while the latter provides a closer look at performance during the financial crisis.  We see that the fitted values lag GDP slightly less and track turning points somewhat more quickly. Table \ref{tab:pvaluedhp} shows p-values for tests of the joint null hypothesis that all coefficients $\hat{\beta} = 0$. We see that the use of filtered indicators improves their statistical significance at almost all forecast horizons, particularly for $h > 2$ quarters. 
%Figures \ref{direct_hp_forecasts} (all observations with exclusion of the pandemic) and \ref{direct_hp_forecasts_financial_crisis} (financial crisis). A comparison with the classic (unfiltered) direct forecasts in the latter figure %in Fig.\eqref{direct_wc_financial_crisis} 
% indicates that the new predictors are slightly less retarded and track the turning points more promptly (`faster'). 
% The filtered approach also exhibits superior statistical properties. As shown in Table \ref{tab:f_stats}, 
%The p-values of the F-statistics for the HP-based forecasts are smaller than for the direct forecasts specifically at $h=3$, but the predictor remains statistically insignificant at forecast horizons exceeding two quarters (even though the evaluation is conducted in-sample and based on all available indicators).

\begin{figure}[h]
    \begin{center}
        \includegraphics[width=0.85\textwidth]{./Figures/direct_hp_forecasts.pdf}
        \caption{Direct HP forecasts: entire data set with exclusion of the pandemic.
        \label{direct_hp_forecasts}}
    \end{center}
\end{figure}

\begin{figure}[h]
    \begin{center}
        \includegraphics[height=5in, width=6in]{./Figures/direct_hp_forecasts_financial_crisis.pdf}
        \caption{Shifted GDP (black), direct HP forecast (red) and classic direct forecast (blue) over the curse of the financial crisis: regression based on all indicators and full sample information (without pandemic).
        \label{direct_hp_forecasts_financial_crisis}}
    \end{center}
\end{figure}

@@

%\begin{table}[ht]
%\centering
%\begin{tabular}{rrrrr}
%  \hline
% & h=3 & h=4 & h=5 & h=6 \\ 
%  \hline
%rRMSE & 0.962 & 0.967 & 0.990 & 0.967 \\ 
%   \hline
%\end{tabular}
%\caption{Relative root mean-square error of HP-C against classic direct forecast for $h\geq %3$.\\Values $<1$ imply  HP-C has smaller RMSE.  } 
%\label{tab:rRMSE}
%\end{table}
\begin{table}[h]
\centering
\begin{tabular}{rrrrrrrr}
  \hline
 & h= 0 & h= 1 & h= 2 & h= 3 & h= 4 & h= 5 & h= 6 \\ 
  \hline
  Unfiltered & 0.028 & 0.000 & 0.001 & 0.910 & 0.509 & 0.091 & 0.178 \\ 
HP-C filtered & 0.000 & 0.001 & 0.036 & 0.046 & 0.049 & 0.012 & 0.022 \\ 
   \hline
\end{tabular}
\caption{p-values for $H_0: \hat{\beta} = 0$, HAC-adjusted Wald test.\\Estimation over full sample (without pandemic.)} 
\label{tab:pvaluedhp}
\end{table}

@@

Revisions refer to changes in the historical values of a time series resulting from updates with new information. Revisions of the M-SSA component predictor can arise from data updates, which we do not analyze further here\footnote{Except for GDP, the other indicators are not or only slightly revised. Moreover, the smoothing effect of filters mitigates the effect of revisions when compared to direct forecasts.}, or from updates to predictor parameters. The latter can be further separated into changes in VAR parameters and regression weights.

>>>>

\usepackage{graphicx} % Required for inserting images
\usepackage{tabularx}
\usepackage{natbib}
\usepackage[spaces,hyphens]{url}
\usepackage[hidelinks]{hyperref}
\usepackage{setspace}
\usepackage{booktabs}
\usepackage{hyphenat}
\usepackage{color}
%\usepackage{subfig}
% \usepackage[justification=left,hang]{caption}
\usepackage[justification=raggedrig,hang]{caption}
\usepackage{subcaption}
	%\captionsetup[subfigure]{labelfont=normal,textfont=normalfont,singlelinecheck=off,justification=centering}
\usepackage[table]{xcolor} 
\usepackage{colortbl}
\usepackage[hang]{footmisc}
\usepackage{appendix}
\usepackage{float}

@@

\begin{abstract} %max.
This paper addresses the trade-off between informational efficiency and forecast smoothness in economic forecasting. Efficient forecasts are often volatile, limiting their practical usefulness. We propose a novel method that employs the framework of Multivariate Smooth Sign Accuracy (M-SSA) to extract smoothed components from leading indicators. This approach enhances the signal-to-noise ratio and allows for systematic control over forecast volatility and timing. Applied to German GDP, our method yields smoothed forecasts that improve the accuracy of traditional models, particularly over medium-term horizons. While smoother forecasts may lag slightly, this can be mitigated by adjusting the forecast horizon. Our findings offer practical tools for forecasters and policymakers subject to adjustment costs or frictions in updating forecasts or implementing policy changes.

@@

\begin{quote}
     \quad ``As we parse the incoming information, we are focused on separating the signal from the noise as the outlook evolves.''\\
    \textit{Speech by Federal Reserve Board Chair Jerome H. Powell, March 07, 2025, \\
    University of Chicago Booth School of Business}
\end{quote}

@@

This study proposes a methodological innovation that extends the classical direct forecasting framework by integrating dimension-reduced information extracted through Multivariate Smooth Sign Accuracy (M-SSA). The core idea is to replace raw
%, unprocessed 
leading indicators with smoothed and filtered components derived from the M-SSA procedure, thereby improving the signal-to-noise ratio in the predictor variables used to forecast GDP. We apply this framework to well-established German economic indicators to show its ability to improve forecasts of GDP and GDP trends using a variety of sample periods and performance metrics.

Surprisingly, we find that smoothing indicator-based forecasts often provides useful leading information for GDP, especially over mid-term horizons ranging from two to four quarters ahead. We also show that while smoother GDP forecasts lag efficient MSE forecasts slightly, this can be mitigated by adjustments in the forecast horizon.\footnote{The lag reflects the trilemma highlighted by \cite{McElroy2019} in which forecasters choose between forecast accuracy, timeliness and volatility.}

@@

Section \ref{sec:data} describes our data series, the sample period used, and documents their dynamic behaviour. Section \ref{sec:direct_forecast} provides benchmark forecasts using simple direct projections of GDP based on our indicators. We demonstrate how in-sample forecast performance may improve when we ``de-noise'' the indicators prior to projection with simple smoothing filters. Section \ref{sec:mSSA} defines the M-SSA method and shows how its forecast performance compares to that of direct projections as well as how it varies with forecast horizon. Section \ref{sec:mSSA_component} considers the ability of our indicators to forecast German GDP growth using the M-SSA framework, and Section \ref{sec:revisions} investigates the temporal stability of the estimated forecasting equations. Section \ref{sec:conclusions} summarizes the results and concludes. The Appendix provides some additional results.

\section{Data and Dependence}\label{sec:data}
\subsection{Data}
The analysis employs a set of well-established leading indicators to forecast quarterly German GDP \citep{Drechsel2012financial,Heinisch2018bottom}, including industrial production (IP), the ifo Business Climate Index (ifo\_c), the Economic Sentiment Indicator (ESI), and the term spread between 10-year bond yields and the 3-month EURIBOR rate (spre\_10y\_3m) covering the period 1995 to 2024.  
We take into account the publication lag of the indicators and the GDP series.\footnote{Data revisions are not considered, given that previous studies \citep{Heinisch2019} have shown that their overall effect on forecast performance is minor.} 
While GDP is available with a one-quarter delay (published in mid of the next quarter), and IP is released with a one-month lag, business surveys and financial data are published without any lag, see Table \ref{tab:data_stucture} which presents the data structure as of the end of January 2025, just before the publication of the Flash estimate for the previous quarter.\footnote{Starting in the second quarter of 2020, the German Federal Statistical Office published for the first time the gross domestic product (GDP) 30 days after the end of the quarter.} All series are consequently re-aligned to correspond with the information available to forecasters, with GDP and IP variables shifted forward according to their respective lags. Table \ref{tab:data_stucture_aligned} displays the dataset as of the end of 2024, with GDP and IP variables temporally aligned at the sample endpoint. This configuration reflects the matrix of explanatory variables accessible in Q4-2024 for the purposes of nowcasting or forecasting GDP. For illustrative purposes, an additional target column is included on the right side of the table, representing the dependent variable for nowcasting applications. In the case of an $h$-step ahead forecast, this column is shifted upwards by $h$ quarters.

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
  \hline
 & GDP & ip & ifo\_c & ESI & spr\_10y\_3m \\ 
  \hline
2024-07-01 &  & 90.700 & 85.800 & 92.500 & -1.300 \\ 
  2024-08-01 &  & 93.200 & 84.900 & 90.900 & -1.400 \\ 
  2024-09-01 &902.571  & 91.200 & 83.700 & 89.800 & -1.400 \\ 
  2024-10-01 &  & 90.800 & 84.100 & 90.700 & -1.200 \\ 
  2024-11-01 &  & 92.200 & 83.800 & 89.300 & -0.900 \\ 
  2024-12-01 &  &  & 82.800 & 86.900 & -0.800 \\ 
  2025-01-01 &  &  & 82.500 & 88.100 & -0.200 \\ 
   \hline
\end{tabular}
\caption{Ragged end as of January 2025, before publication of the flash estimate.} 
\label{tab:data_stucture}
\end{table}

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrrc}
  \hline
 & GDP & ip & ifo\_c & ESI & spr\_10y\_3m &&Target (nowcast)\\ 
  \hline
  2024-07-01 & & 93.400 & 85.800 & 92.500 & -1.300&|& \\ 
  2024-08-01 & & 90.700&  84.900 & 90.900 & -1.400&| &\\ 
  2024-09-01 & 901.623 &93.200 & 83.700 & 89.800 & -1.400&| &902.571\\ 
  2024-10-01 &  &91.200 & 84.100 & 90.700 & -1.200&| &\\ 
  2024-11-01 &  & 90.800& 83.800 & 89.300 & -0.900&| &\\ 
  2024-12-01 &902.571   & 92.200  & 82.800 & 86.900 & -0.800&| &?\\ 
    \hline
\end{tabular}
\caption{Columns 1 through 5 constitute the matrix of explanatory variables available in Q4-2024 for the establishment of a GDP nowcast or forecast. All series are artificially aligned at the sample endpoint. Column 6 represents the target variable for the nowcast in Q4-2024.} 
\label{tab:data_stucture_aligned}
\end{table}

To ensure comparability and mitigate scale effects, all indicators (except the spread, which is differenced only) are log-differenced and standardized, the latter applied solely for ease of visual inspection. Additionally, extreme outliers associated with the COVID-19 pandemic are trimmed to prevent distortions in subsequent figures. 
 Monthly series are transformed to quarterly series and are displayed in Figure \ref{fig:data}, while Figure \ref{fig:data_lags} offers a close-up view of the financial crisis and the pandemic.\footnote{Monthly series are averaged to quarterly values.} The latter figure illustrates that the real-time GDP and industrial production series are synchronized at dips and peaks (coincident), whereas the other indicators tend to lead somewhat. The latter can be leveraged in a multivariate approach to enhance forecasting performance compared to univariate benchmarks.

@@

@@

Figures \ref{CCF} (full sample) and \ref{CCF_wc} (omitting the Pandemic period) show the sample cross-correlation functions (CCF's) between current and lagged GDP and the two survey indicators. Both figures show strong correlations that peak at leads relative to GDP. Results for IP showed no lead, while correlations for the term spread were weaker. The figures also show that the precise strengths of the correlations are influenced by the outliers associated with the COVID-19 pandemic. In the interest of model stability, we exclude data from Q4-2019 to Q4-2020 (pandemic) when modelling the dynamics for the multivariate filters developed below.

@@

We then estimated VARMA models to capture these dynamics, selecting a parsimonious VAR(1) specification.\footnote{We used the \texttt{refVARMA()} procedure in the \cite{MTSpackage} \texttt{MTS} package, which sets VARMA coefficients with t-statistics $< 1.5$ to zero \citep[see][]{tsay2013multivariate}. We intend to explore richer dynamic structures in subsequent work.} Figure \ref{ma_inv_BIP} shows the implied responses of GDP to shocks to each indicator variable.\footnote{Shown are the MA coefficients for GDP from the vector moving-average (VMA) model obtained by inversion of the VAR(1) representation (i.e. the impluse-response functions, or IRFs, of GDP to the correlated reduced-form shocks.)}

\begin{figure}[h]
    \begin{center}
        \includegraphics[height=3in, width=6in]{./Figures/ma_inv_BIP.pdf}
        \caption{VMA coefficients for GDP Implied by VAR(1)\\Estimation with pandemic (left) and without pandemic (right).
        \label{ma_inv_BIP}}\end{center}
\end{figure}

The model summarizes key features of the data after we account for publication lags. GDP responds most strongly to shocks to the two survey indicators (ifo\_c and ESI) after a modest lag. There also appears to be a more modest response to IP shocks after a slightly longer lag, but this vanishes once data after the start of the pandemic is included.\footnote{There is also weaker evidence of a response to the term spread (spr\_10y\_3m) at longer lags. } The delayed response of GDP to movements in these various indicators suggests that multivariate models may provide better GDP forecasts than univariate models, a question we pursue in the next section.
%, where lagging series (GDP and IP, after accounting for publication lags) depend on indicators leading in relative terms. In particular, the impulse response of the spread suggests the relevance of a univariate model for this specific leading indicator.

@@

\begin{equation}
    u_{t+h} = \beta' \cdot \mathbf{X}_t + \varepsilon_t
    \label{eq:direct_forecast}
\end{equation}

where
\begin{itemize}
    \item $u_{t+h}$ is the growth rate of GDP, $h = \{0, 1, 2, ... \}$
    \item $\mathbf{X}_t \equiv [1,u_{t-1}, ip_t, ifo\_c_t, ESI_t, spr\_10y\_3m_t]'$ or their transformed counterparts and $u_{t-1}$ accounts for the publication lag of GDP.\footnote{Including lagged GDP enables the nesting of the AR forecast within the broader modeling framework.}
    \item $\hat{\beta}$ is estimated by least-squares, omitting observations affected by the pandemic, i.e. Q4-2019 to Q4-2020.
\end{itemize}

@@

\begin{figure}[H]
    \begin{center}
        \includegraphics[width=0.9\textwidth]{./Figures/direct_wc_all.pdf}
        %\includegraphics[height=3in, width=4.5in]{./Figures/direct_wc_all.pdf}
        \caption{Direct forecasts (red lines) based on all indicators and full sample information (without pandemic): GDP (black line) shifted (i.e. lagged) by zero, one, two, and three quarters.
        \label{direct_wc}}
    \end{center}
\end{figure}

Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} compare actual and fitted values for GDP at forecast horizons $h = 0, \ldots, 3$, using the four indicators and lagged GDP, without applying any filtering. Figure \ref{direct_wc} shows results over the sample without the pandemic, while Figure \ref{direct_wc_financial_crisis} provides a closer look at the fit during the financial crisis. Table \ref{tab:pvaluedhp} shows p-values for tests of the joint null hypothesis that all coefficients on the indicators are zero (HAC-adjusted Wald test).

\begin{figure}[H]
    \begin{center}
        \includegraphics[width=0.9\textwidth]{./Figures/direct_wc_financial_crisis.pdf}
        \caption{Direct forecast across the financial crisis:  regression based on all indicators and full sample information, without pandemic.
        \label{direct_wc_financial_crisis}}
    \end{center}
\end{figure}

The results indicate that as the forecast horizon $h$ increases, the turning points (i.e., peaks and troughs) of the direct forecasts become increasingly right-shifted (lagged) relative to the target GDP. Additionally, the p-values of the HAC-adjusted Wald test precipitously increase from $h=2$ to $h=3$ (see Table \ref{tab:pvaluedhp}), implying that the indicators become statistically insignificant for forecast horizons of more than two quarters.\footnote{ 
These findings are confirmed by the relative root mean-square errors comparing direct forecasts against the mean benchmark, see Table \ref{rRMSE_mSSA_direct_mean_without_covid8} in the Appendix.}

@@

Figure \ref{hp_160} presents both classic two-sided and one-sided HP(160) (denoted HP-C for \textit{concurrent}) filters (top panels) along with their associated amplitude functions in the frequency domain (bottom panels). It confirms that the HP-C filter assigns greater weight to yearly components—relevant in a mid-term forecast context— than the classic quarterly HP filter, while also reducing undesirable high-frequency noise. Therefore, while the choice of the HP(160) filter in this example
is somewhat arbitrary, it should serve the purpose of reducing high-frequency noise while preserving longer-duration movements.\footnote{A comprehensive technical analysis of the effect of $\lambda$ on the resulting GDP predictor can be conducted using the M-SSA package (\url{https://github.com/wiaidp/R-package-SSA-Predictor}).}

@@

Figures \ref{direct_hp_forecasts} and \ref{direct_hp_forecasts_financial_crisis} provide results comparable to those of Figures \ref{direct_wc} and \ref{direct_wc_financial_crisis} but now using indicators smoothed with the HP-C filter, covering the full sample without the pandemic and providing a closer look at performance during the financial crisis.  The fitted values lag GDP slightly less and track turning points somewhat more quickly. Table \ref{tab:pvaluedhp} shows p-values for tests of the joint null hypothesis that all coefficients $\hat{\beta} = 0$. The use of filtered indicators improves their statistical significance at forecast horizons $h > 2$ quarters. 
%Figures \ref{direct_hp_forecasts} (all observations with exclusion of the pandemic) and \ref{direct_hp_forecasts_financial_crisis} (financial crisis). A comparison with the classic (unfiltered) direct forecasts in the latter figure %in Fig.\eqref{direct_wc_financial_crisis} 
% indicates that the new predictors are slightly less retarded and track the turning points more promptly (`faster'). 
% The filtered approach also exhibits superior statistical properties. As shown in Table \ref{tab:f_stats}, 
%The p-values of the F-statistics for the HP-based forecasts are smaller than for the direct forecasts specifically at $h=3$, but the predictor remains statistically insignificant at forecast horizons exceeding two quarters (even though the evaluation is conducted in-sample and based on all available indicators).

\begin{figure}[H]
    \begin{center}
        \includegraphics[width=0.85\textwidth]{./Figures/direct_hp_forecasts.pdf}
        \caption{Direct HP forecasts: entire data set with exclusion of the pandemic.
        \label{direct_hp_forecasts}}
    \end{center}
\end{figure}

\begin{figure}[H]
    \begin{center}
        \includegraphics[height=5in, width=6in]{./Figures/direct_hp_forecasts_financial_crisis.pdf}
        \caption{Shifted GDP (black), direct HP forecast (red) and classic direct forecast (blue) over the course of the financial crisis: regression based on all indicators and full sample information (without pandemic).
        \label{direct_hp_forecasts_financial_crisis}}
    \end{center}
\end{figure}

@@

%\begin{table}[ht]
%\centering
%\begin{tabular}{rrrrr}
%  \hline
% & h=3 & h=4 & h=5 & h=6 \\ 
%  \hline
%rRMSE & 0.962 & 0.967 & 0.990 & 0.967 \\ 
%   \hline
%\end{tabular}
%\caption{Relative root mean-square error of HP-C against classic direct forecast for $h\geq %3$.\\Values $<1$ imply  HP-C has smaller RMSE.  } 
%\label{tab:rRMSE}
%\end{table}
\begin{table}[h]
\centering
\begin{tabular}{lrrrrrr}
  \hline
  & h= 1 & h= 2 & h= 3 & h= 4 & h= 5 & h= 6 \\ 
  \hline
  Unfiltered & 0.000 & 0.001 & 0.910 & 0.509 & 0.091 & 0.178 \\ 
HP-C filtered & 0.001 & 0.036 & 0.046 & 0.049 & 0.012 & 0.022 \\ 
   \hline
\end{tabular}
\caption{p-values for $H_0: \hat{\beta} = 0$, HAC-adjusted Wald test.\\Estimation over full sample (without pandemic.)} 
\label{tab:pvaluedhp}
\end{table}

@@

Revisions refer to changes in the historical values of a time series resulting from updates with new information. Revisions of the M-SSA component predictor can arise from data updates, which we do not analyze further here\footnote{Except for GDP, the other indicators are not or only slightly revised. Moreover, the smoothing effect of filters mitigates the effect of revisions when compared to direct forecasts.}, or from updates to predictor parameters. The latter can be further separated into changes in the regression weights and VAR parameters.
